---
title: "Timeseries 101"
output: output: 
  html_document:
    keep_md: true

---

```{r, echo = FALSE}

knit(input="readme.rmd", output = "readme.md") #see ?knit for more options

```


# Summarof Material from Rob Hyndman's Timeseries Forecasting

More in depth explanations can be located here in his textbook:
https://otexts.com/fpp2/


```{r install packages, message=FALSE, warning=FALSE}

library(fpp2)
library(dplyr)

```


## Timeseries Analysis Examples

### Basic Plots, outliers, and frequency

```{r}

# Plot the three series
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)

# Find the outlier in the gold series
goldoutlier <- which.max(gold)
goldoutlier

# Look at the seasonal frequencies of the three series
frequency(gold)
frequency(woolyrnq)
frequency(gas)

```


### Basic TS object creation, autoplot, subseries plot

```{r}

#create basic timeseries from data frame
# insert times sequence and data - TO DO
#myts <- ts(mydata[,c(2:4)], start = c(1981, 1), frequency = 4)

# beer data
beer = ausbeer
beer

# Make plots of the beer data
autoplot(beer)
ggsubseriesplot(beer)

```


### Polar plots, more suberies

```{r}

# Create plots of the a10 data
autoplot(a10)
ggseasonplot(a10)

# Produce a polar coordinate season plot for the a10 data
ggseasonplot(a10, polar = TRUE)

# Restrict the ausbeer data to start in 1992
beer <- window(ausbeer, start = 1992)

# Make plots of the beer data
autoplot(beer)
ggsubseriesplot(beer)

```

## Intro to ACF Plots

```{r}

# Create an autoplot of the oil data
autoplot(oil)

# Create a lag plot of the oil data
gglagplot(oil)

# Create an ACF plot of the oil data
ggAcf(oil)

```


### Find max autocorrelation with ACF plots

```{r}


# Plot the annual sunspot numbers
autoplot(sunspot.year)
ggAcf(sunspot.year)

# Save the lag corresponding to maximum autocorrelation
maxlag_sunspot <- 1

# Plot the traffic on the Hyndsight blog
autoplot(hyndsight)
ggAcf(hyndsight)

# Save the lag corresponding to maximum autocorrelation
maxlag_hyndsight <- 7


```


### Plot differentiated series, ACF plot of differentiated series, Ljung Box Test
```{r}

# Plot the original series
autoplot(goog)

# Plot the differenced series
autoplot(diff(goog))

# ACF of the differenced series
ggAcf(diff(goog))

# Ljung-Box test of the differenced series
Box.test(diff(goog), lag = 10, type = "Ljung")

```


## Naive Forecast, Seasonal Naive Forecast, Error Measures (RMSE)

The very simplest forecasting method is to use the most recent observation; this is called a naive forecast and can be implemented in a namesake function. This is the best that can be done for many time series including most stock price data, and even if it is not a good forecasting method, it provides a useful benchmark for other forecasting methods.

For seasonal data, a related idea is to use the corresponding season from the last year of data. For example, if you want to forecast the sales volume for next March, you would use the sales volume from the previous March. This is implemented in the snaive() function, meaning, seasonal naive.

```{r}

# Use naive() to forecast the goog series
fcgoog <- naive(goog, h = 20)

# Plot and summarize the forecasts
autoplot(fcgoog)
summary(fcgoog)

# Use snaive() to forecast the ausbeer series
fcbeer <- snaive(ausbeer, h = 16)

# Plot and summarize the forecasts
autoplot(fcbeer)
summary(fcbeer)

```
### Residuals, checkresidulas()

When applying a forecasting method, it is important to always check that the residuals are well-behaved (i.e., no outliers or patterns) and resemble white noise. The prediction intervals are computed assuming that the residuals are also normally distributed. You can use the checkresiduals() function to verify these characteristics; it will give the results of a Ljung-Box test.


```{r}

# Check the residuals from the naive forecasts applied to the goog series
goog %>% naive() %>% checkresiduals()

# Do they look like white noise (TRUE or FALSE)
googwn <- TRUE


# Check the residuals from the seasonal naive forecasts applied to the ausbeer series
ausbeer %>% snaive() %>% checkresiduals()

# Do they look like white noise (TRUE or FALSE)
beerwn <- FALSE

```

### Subset training and test sets

```{r}

# Create the training data as train
train <- subset(gold, end = 1000)

# Compute naive forecasts and save to naive_fc
naive_fc <- naive(train, h = 108)

# Compute mean forecasts and save to mean_fc
mean_fc <- meanf(train, h = 108)

# Use accuracy() to compute RMSE statistics
accuracy(naive_fc, gold)
accuracy(mean_fc, gold)

# Assign one of the two forecasts as bestforecasts
bestforecasts <- naive_fc

```

### Cross Validation

```{r}

# Compute cross-validated errors for up to 8 steps ahead
e <- tsCV(goog, forecastfunction = naive, h = 8)

# Compute the MSE values and remove missing values
mse <- colMeans(e^2, na.rm = TRUE)

# Plot the MSE values against the forecast horizon
data.frame(h = 1:8, MSE = mse) %>%
  ggplot(aes(x = h, y = MSE)) + geom_point()

```
## Exponential Smoothing

```{r}

# Use ses() to forecast the next 10 years of winning times
fc <- ses(marathon, h = 10)

# Use summary() to see the model parameters
summary(fc)

# Use autoplot() to plot the forecasts
autoplot(fc)

# Add the one-step forecasts for the training data to the plot
autoplot(fc) + autolayer(fitted(fc))


```

### Compare Exponential Smoothing vs Naive using RMSE

```{r}

# Create a training set using subset()
train <- subset(marathon, end = length(marathon) - 20)

# Compute SES and naive forecasts, save to fcses and fcnaive
fcses <- ses(train, h = 20)
fcnaive <- naive(train, h = 20)

# Calculate forecast accuracy measures
accuracy(fcses, marathon)
accuracy(fcnaive, marathon)

# Save the best forecasts as fcbest
fcbest <- fcnaive




```

## Holt's Method

```{r}

# Produce 10 year forecasts of austa using holt()
fcholt <- holt(austa, h = 10)

# Look at fitted model using summary()
summary(fcholt)

# Plot the forecasts
autoplot(fcholt)

# Check that the residuals look like white noise
checkresiduals(fcholt)


```
### Holt Winter's 3 year monthly forecast (multiplicative method)

```{r}


# Plot the data
autoplot(a10)

# Produce 3 year forecasts
fc <- hw(a10, seasonal = "multiplicative", h = 36)

# Check if residuals look like white noise
checkresiduals(fc)
whitenoise <- FALSE

# Plot forecasts
autoplot(fc)






```
### Holt Winter's Additive Forecast vs Seasonal Naive

```{r}


# Create training data with subset()
train <- subset(hyndsight, end = length(hyndsight) - 28)


# Holt-Winters additive forecasts as fchw
fchw <- hw(train, seasonal = "additive", h = 28)

# Seasonal naive forecasts as fcsn
fcsn <- snaive(train, h = 28)

# Find better forecasts with accuracy()
accuracy(fchw, hyndsight)
accuracy(fcsn, hyndsight)

# Plot the better forecasts
autoplot(fchw)




```
## ETS

```{r}

# Fit ETS model to austa in fitaus
fitaus <- ets(austa)

# Check residuals
checkresiduals(fitaus)

# Plot forecasts
autoplot(forecast(fitaus))

# Repeat for hyndsight data in fiths
fiths <- ets(hyndsight)
checkresiduals(fiths)
autoplot(forecast(fiths))

# Which model(s) fails test? (TRUE or FALSE)
fitausfail <- FALSE
fithsfail <- TRUE



```
### ETS vs Seasonal Naive

```{r}
# Function to return ETS forecasts
fets <- function(y, h) {
  forecast(ets(y), h = h)
}

# Apply tsCV() for both methods
e1 <- tsCV(cement, fets, h = 4)
e2 <- tsCV(cement, snaive, h = 4)

# Compute MSE of resulting errors (watch out for missing values)
mean(e1^2, na.rm = TRUE)
mean(e2^2, na.rm = TRUE)

# Copy the best forecast MSE
bestmse <- mean(e2^2, na.rm = TRUE)


```

### ETS 20 year forecast on Lynx data set

```{r}

# Plot the lynx series
autoplot(lynx)

# Use ets() to model the lynx series
fit <- ets(lynx)

# Use summary() to look at model and parameters
summary(fit)

# Plot 20-year forecasts of the lynx series
fit %>% forecast(h=20) %>% autoplot()





```

## Box Cox Transformations - compare lambdas 

```{r}

# Plot the series
autoplot(a10)

# Try four values of lambda in Box-Cox transformations
a10 %>% BoxCox(lambda = 0.0) %>% autoplot()
a10 %>% BoxCox(lambda = 0.1) %>% autoplot()
a10 %>% BoxCox(lambda = 0.2) %>% autoplot()
a10 %>% BoxCox(lambda = 0.3) %>% autoplot()

# Compare with BoxCox.lambda()
a10 %>% BoxCox.lambda()



```
## Diff and ACF Diff plots

```{r}


# Plot the US female murder rate
autoplot(wmurders)

# Plot the differenced murder rate
autoplot(diff(wmurders))

# Plot the ACF of the differenced murder rate
ggAcf(diff(wmurders))



```


## Log and seasonal differnce to make time series stationary

```{r}

# Plot the data
autoplot(h02)

# Take logs and seasonal differences of h02
difflogh02 <- diff(log(h02), lag = 12)

# Plot difflogh02
autoplot(difflogh02)

# Take another difference and plot
ddifflogh02 <- diff(difflogh02)
autoplot(ddifflogh02)

# Plot ACF of ddifflogh02
ggAcf(ddifflogh02)


```


## ARIMA


```{r}

# Fit an automatic ARIMA model to the austa series
fit <- auto.arima(austa)

# Check that the residuals look like white noise
checkresiduals(fit)
residualsok <- TRUE ## check p value ensuring its > 0.05 (white noise)

# Summarize the model
summary(fit)


# Find the AICc value and the number of differences used
AICc <- -14.46
d <- 1

# Plot forecasts of fit
fit %>% forecast(h = 10) %>% autoplot()


```
### ARIMA - detailed

The Arima() function can be used to select a specific ARIMA model. Its first argument, order, is set to a vector that specifies the values of ,  and . The second argument, include.constant, is a booolean that determines if the constant , or drift, should be included. 

```{r}

# Plot forecasts from an ARIMA(0,1,1) model with no drift
austa %>% Arima(order = c(0, 1, 1), include.constant = FALSE) %>% forecast() %>% autoplot()

# Plot forecasts from an ARIMA(2,1,3) model with drift
austa %>% Arima(order = c(2, 1, 3), include.constant = TRUE) %>% forecast() %>% autoplot()
# Plot forecasts from an ARIMA(0,0,1) model with a constant
austa %>% Arima(order = c(0, 0, 1), include.constant = TRUE) %>% forecast() %>% autoplot()
# Plot forecasts from an ARIMA(0,2,1) model with no constant
austa %>% Arima(order = c(0, 2, 1), include.constant = FALSE) %>% forecast() %>% autoplot()


```

### ETS vs ARIMA on non seasonal data

```{r}

# Set up forecast functions for ETS and ARIMA models
fets <- function(x, h) {
  forecast(ets(x), h = h)
}
farima <- function(x, h) {
  forecast(auto.arima(x), h = h)
}

# Compute CV errors for ETS on austa as e1
e1 <- tsCV(austa, fets, h = 1)

# Compute CV errors for ARIMA on austa as e2
e2 <- tsCV(austa, farima, h = 1)

# Find MSE of each model class
mean(e1^2, na.rm = TRUE)
mean(e2^2, na.rm = TRUE)

# Plot 10-year forecasts using the best model class
austa %>% farima(h = 10) %>% autoplot()

```
### Seasonal ARIMA Forecast

```{r}

# Check that the logged h02 data have stable variance
h02 %>% log() %>% autoplot()

# Fit a seasonal ARIMA model to h02 with lambda = 0
fit <- auto.arima(h02, lambda = 0)

# Summarize the fitted model
summary(fit)

# Record the amount of lag-1 differencing and seasonal differencing used
d <- 1
D <- 1

# Plot 2-year forecasts
fit %>% forecast(h = 24) %>% autoplot()








```
### Deep diving auto.arima()

Exploring auto.arima() options
The auto.arima() function needs to estimate a lot of different models, and various short-cuts are used to try to make the function as fast as possible. This can cause a model to be returned which does not actually have the smallest AICc value. To make auto.arima() work harder to find a good model, add the optional argument stepwise = FALSE to look at a much larger collection of models.

```{r}

# Find an ARIMA model for euretail
fit1 <- auto.arima(euretail)

# Don't use a stepwise search
fit2 <- auto.arima(euretail, stepwise = FALSE)

# Compare Models to decide which to deploy

#summary(fit1)
#summary(fit2)

# AICc of better model
AICc <- 68.39

# Compute 2-year forecasts from better model
fit2 %>% forecast(h=8) %>% autoplot()


```


### ARIMA vs ETS

```{r}

# Use 20 years of the qcement data beginning in 1988
train <- window(qcement, start = 1988, end = c(2007, 4))

# Fit an ARIMA and an ETS model to the training data
fit1 <- auto.arima(train)
fit2 <- ets(train)

# Check that both models have white noise residuals
checkresiduals(fit1)
checkresiduals(fit2)

# Produce forecasts for each model
fc1 <- forecast(fit1, h = 25)
fc2 <- forecast(fit2, h = 25)

# Use accuracy() to find best model based on RMSE
accuracy(fc1, qcement)
accuracy(fc2, qcement)
bettermodel <- fit2


```

## ARIMA + Regressors

```{r}

# Time plot of both variables
autoplot(advert, facets = TRUE)

# Fit ARIMA model
fit <- auto.arima(advert[, "sales"], xreg = advert[, "advert"], stationary = TRUE)

# Check model. Increase in sales for each unit increase in advertising
salesincrease <- coefficients(fit)[3]

# Forecast fit as fc
fc <- forecast(fit, xreg = rep(10, 6))

# Plot fc with x and y labels
autoplot(fc) + xlab("Month") + ylab("Sales")




```


```